---
title: 'Statistical Inference: Project 1 Part 1'
author: "JR Sturino"
date: "November 17, 2015"
output: word_document
---

##Overview
The task is to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations.

##Simulations
Run 1000 simulations using the parameters in the overview section. The following code loops through each of our simulations and obtains the mean:

```{r}
startSimulations = 1
endSimulations = 1000
lambda <- 0.2
n <- 40

set.seed(1)
simulationMeans <- sapply(startSimulations:endSimulations, function(x) {mean(rexp(n, lambda))})
head(simulationMeans)
```

##Sample Mean versus Theoretical Mean

###Sample Mean
Obtain the mean for our data set from the simulation
```{r}
mean(simulationMeans)
```

###Theoretical Mean
We calculate the theoretical mean using 1/lambda
```{r}
(1/lambda)
```

###Conclusion
In this case the sample and theoretical mean are quite similar. 

##Sample Variance versus Theoretical Variance

###Sample Variance
Obtain the variance from the data set, we do this by obtaining the standard deviation and squaring it:
```{r}
sd(simulationMeans)^2
```

###Theoretical Variance
We obtain the theoretical variance by using 1/lambda/sqrt(n) and squaring the result:
```{r}
((1/lambda)/(sqrt(n)))^2
```

###Conclusion
Like the mean, the variance is also very close.

##Distribution
In the below graph we plot two lines vertically; the red line plots the theoretical mean while the green plots the sample mean.


```{r message=FALSE}
library(ggplot2)
meanData <- data.frame(x=simulationMeans)
tVar <- ((1/lambda)/(sqrt(n)))^2

ggplot(meanData, aes(x = simulationMeans)) + 
  geom_histogram(aes(y=..density..), color="black",fill="white") + 
  xlab("Mean of Simulation") + 
  ylab("Mean Density") + 
  geom_vline(xintercept=mean(meanData$x), color="green", lwd=1) +
  geom_vline(xintercept=(1/lambda), color="red", lwd=1) +
  stat_function(fun=dnorm, args=list(mean=mean(meanData$x), sd=sd(meanData$x)^2), color="blue", lwd=1.5) +
  stat_function(fun=dnorm, args=list(mean=1/lambda, sd=tVar), color="yellow", lwd=1.5)
```

Looking at the graph, the blue line represents the sample distrubtion and the yellow line represents the theoretical distrubtion. As you can see they both follow the normal distrubtion.
